Detection of slow slip events using wavelet analysis of GNSS recordings

### 2021-10-21 paper/paper_bssa.tex
Reformat paper for BSSA

%% March 2018
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
% There are two options for article format:
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%

%% To submit your paper:
\documentclass[draft]{agujournal2018}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Enter journal name here}

@@

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{Detection of slow slip events using wavelet analysis of GNSS recordings}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{A. Ducellier \affil{1}, K. C. Creager \affil{1}, D. A. Schmidt \affil{1}}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{1}{University of Washington}
%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Ariane Ducellier}{ducela@uw.edu}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 100 characters or less with no special characters or punctuation

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 100 characters or less with no special characters or punctuation
% \end{keypoints}

\begin{keypoints}
\item enter point 1 here
\item enter point 2 here
\item enter point 3 here
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
% A good abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}
Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks. Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles. In many places, tectonic tremor is also observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in GNSS noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted averages of a time series vary from one averaging period to the next. In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a peak in the wavelet signal. We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}).
\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

@@

In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8. They also observed many smaller bursts of tremor of duration 1 to 50 hours in between the big ETS events, without any detectable signal in the GPS data. However, based on the relationship between slow slip moment and number of hours of tremor for bigger events, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with the tremor bursts. This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}). \\

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.pdf}
\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.pdf}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.pdf}
\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.pdf}

 \begin{table}
     \caption{Big Episodic Tremor and Slip events. The duration and the number of tremor are from the tremor catalog of \textcolor{red}{reference?}. The event number and the magnitude are from the slow slip catalog of ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c c}
 \hline
 Time & Duration (days) & Number of tremor (hours) & Event number & Magnitude \\
 \hline
 2007.06 & 28 & 398 & 3 & 6.68 \\
 2008.36 & 25 & 402 & 10 & 6.56 \\
 2009.35 & 24 & 248 & 16 & 6.49 \\
 2010.63 & 29 & 518 & 24 & 6.54 \\
 2011.60 & 37 & 479 & 30 & 6.47 \\
 2012.72 & 37 & 620 & 34 & 6.54 \\
 2013.71 & 27 & 423 & 41 & 6.58 \\
 2014.65 & 15 & 190 & 48 & 6.03 \\
 2014.89 & 38 & 385 & 51 & 6.40 \\
 2016.11 & 43 & 421 & 54 & 6.79 \\
 2017.23 & 19 & 279 & 59 & 6.61 \\
 2018.49 & 22 & 381 & & \\
 2019.23 & 34 & 195 & & \\
 2019.88 & 16 & 205 & & \\
 2020.79 & 26 & 193 & & \\
 2020.86 & 12 & 162 & & \\
 2021.09 & 14 & 230 & & \\
 \hline
 \end{tabular}
 \end{table}

@@

@@

 \begin{table}
 \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c}
 \hline
 Time & Event number & Magnitude & Sub-event of bigger event \\
 \hline
 2007.06 & 1 & 5.64 & Yes \\
 2007.08 & 2 & 5.91 & Yes \\
 2008.38 & 11 & 5.50 & Yes \\
 2009.16 & 14 & 5.50 & No \\
 2009.36 & 17 & 5.32 & Yes \\
 2010.63 & 25 & 5.76 & Yes \\
 2011.66 & 31 & 5.61 & Yes \\
 2011.66 & 32 & 5.32 & Yes \\
 2012.69 & 35 & 5.56 & Yes \\
 2013.74 & 42 & 5.71 & Yes \\
 2014.69 & 49 & 5.31 & Yes \\
 2014.93 & 52 & 5.39 & Yes \\
 2016.03 & 57 & 5.80 & Yes \\
 2017.13 & 60 & 5.43 & Yes \\
 2017.22 & 61 & 5.37 & Yes \\
 \hline
 \end{tabular}
 \end{table}

Figure 9 shows the 5th level detail wavelet decomposition of the GPS data. Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.

@@

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}[h]
% \centering
% when using pdflatex, use pdf file:
% \includegraphics[natwidth=800px,natheight=600px]{figsamp.pdf}
%
% when using dvips, use .eps file:
% \includegraphics[natwidth=800px,natheight=600px]{figsamp.eps}
%
% \caption{Short caption}
% \label{figone}
%  \end{figure}
%
% We recommend that you provide the native width and height (natwidth, natheight) of your figures.
% Specifying native dimensions ensures that your figures are properly scaled
%
%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments
This work was funded by the grant from the National Science Foundation XXX. A.D. would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington. The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA}. Figure 4 was done using GMT ~\citep{WES_1991}. The Python scripts used to analyze the data and make the  figures can be found on the  first author's Github account.

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BibTeX is preferred:
%
\bibliography{bibliography.bib}
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).
%% Example \citet and \citep:
%  ...as shown by \citet{Boug10}, \citet{Buiz07}, \citet{Fra10},
%  \citet{Ghel00}, and \citet{Leit74}.

%  ...as shown by \citep{Boug10}, \citep{Buiz07}, \citep{Fra10},
%  \citep{Ghel00, Leit74}.

%  ...has been shown \citep [e.g.,][]{Boug10,Buiz07,Fra10}.

@@

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

>>>>

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{authblk}
\usepackage{color}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{url}

\linenumbers

\title{Detection of slow slip events using wavelet analysis of GNSS recordings}
\author[1]{Ariane Ducellier}
\author[1]{Kenneth C. Creager}
\author[1]{David A. Schmidt}
\affil[1]{University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310}
\date{}

@@

\maketitle

\section*{Key points}

\begin{itemize}
\item We use a wavelet-based signal processing method to detect slow slip events in GNSS data.
\item There is a good correlation between events detected with GNSS data and events detected with seismic data.
\item The method could be applied in regions and no tremor are detected in conjonction with slow slip events.
\end{itemize}

\newpage

\section*{Abstract}

Slow slip events were discovered in many subduction zones during the last two decades thanks to recordings of the displacement of Earth's surface by GNSS networks. Slow slip can last from a few days to several years and has a relatively short recurrence time (months to years), compared to the recurrence time of regular earthquakes (up to several hundreds of years). In many places, tectonic tremor is also observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in GNSS noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. In this paper, we use the Maximal Overlap Discrete Wavelet Transform (MODWT) to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a peak in the wavelet signal. We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only seismic data. The wavelet-based detection method detects all events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citet{MIC_2019}).

@@

In Cascadia and Guerrero, Mexico, tremor has been used as a proxy to observe slow slip events that are not directly detectable in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8. They also observed many smaller bursts of tremor of duration 1 to 50 hours in between the big ETS events, without any detectable signal in the GPS data. However, based on the relationship between slow slip moment and number of hours of tremor for bigger events, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with the tremor bursts. \\ %This leads to a power-law relationship between seismic moment and number of events with a $b$-value close to one, similar to the distribution of normal earthquakes (\textcolor{red}{reference?}). \\

@@

@@

@@

@@

To better evaluate the number of true and false detections, we convert the wavelet details into binary time series. If the absolute value of the wavelet detail is higher that a threshold, we replace the value by 1 (for positive values) or -1 (for negative values), otherwise we replace the value by 0. We do this on both the wavelet details of the GPS data and of the tremor data. Then we decide that if both the GPS and the tremor time series take the value 1 (or both take the value -1), we have a true detection (true positive TP). If the GPS and the tremor time series have opposite signs, or if the absolute value of the GPS time series is 1 but the value of the tremor time series is 0, we have a false detection (false positive FP). If both time series take the value 0, we do not have detection (true negative TN). If the GPS time series take the value 0, but the absolute value of the tremor time series is 1, we miss a detection (false negative). We then define the sensitivity (true positive rate) and the specificity (equal to 1 minus the false positive rate) as:

\begin{equation}
\begin{aligned}
\text{sensitivity} &= \frac{TP}{TP + FN} \\
\text{specificity} &= \frac{TN} {TN + FP}
\end{aligned}
\end{equation}

We can then evaluate the quality of the detections obtained with our method by plotting a receiver operating characteristic curve (ROC curve). The ROC curve is widely use for binary classification problems, and is plotted by varying the values of the threshold (here the thresholds used to convert the GPS and the tremor time series into binary time series), computing the corresponding values of the true positive rate and the false positive rate (equal to 1 minus the specificity), and plotting the true positive rate as a function of the false positive rate. If the classification was made randomly, all the points would fall on the first diagonal. If the classifier was perfect, the corresponding point would fall on the top left corner of the graph with true positive rate equal to 1 and false positive rate equal to 0. The bigger the area under the curve, the better the classifier is. \\

As the slow slip events are better seen on levels 6, 7 and 8 of the wavelet decomposition, we first add the wavelet details corresponding to levels 6 to 8, and transform the resulting time series into a binary time series. We apply this transform to both the GPS and the tremor time series with varying thresholds. We then plot the ROC curve on Figure 8. The corresponding sums of the wavelet details for the GPS data and the tremor data are shown on Figure 10. We can see that there is a trade-off between sensitivity and specificity. If we decrease the false positive rate, we also decrease the number of true events detected. If we increase the number of true events detected, we also increase the false positive rate. \\

@@

Figure 11 shows the 5th level detail wavelet decomposition of the GPS data. Red lines show the timing of the big ETS events from Table 1, and blue lines show the timing of the small slow slip events from Table 2.

@@

\section*{Data and Resources}

The GPS recordings used for this analysis can be downloaded from the PANGA website ~\citep{PANGA} \url{http://www.panga.cwu.edu/}. \textcolor{red}{Reference for tremor catalog}. The Python scripts used to analyze the data and make the figures can be found on the first author's Github account \url{https://github.com/ArianeDucellier/slowslip}. Figure 4 was done using GMT ~\citep{WES_1991}.

\section*{Acknowledgements}

This work was funded by the grant from the National Science Foundation XXX. A.D. would like to thank Pr Donald Percival for introducing her to wavelet methods during his excellent class taught at University of Washington.

\section*{Declaration of Competing Interests}

The authors declare no competing interests.

\bibliographystyle{plainnat}
\bibliography{bibliography}

\newpage

\section*{Addresses}

Ariane Ducellier. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310. \\

Kenneth C. Creager. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310. \\

David A. Schmidt. University of Washington, Department of Earth and Space Sciences, Box 351310, 4000 15th Avenue NE Seattle, WA 98195-1310.

\newpage

\section*{Tables}

\begin{table}[hbt!]
\caption{Big Episodic Tremor and Slip events. The duration and the number of tremor are from the tremor catalog of \textcolor{red}{reference?}. The event number and the magnitude are from the slow slip catalog of ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c c}
 \hline
 Time & Duration (days) & Number of tremor (hours) & Event number & Magnitude \\
 \hline
 2007.06 & 28 & 398 & 3 & 6.68 \\
 2008.36 & 25 & 402 & 10 & 6.56 \\
 2009.35 & 24 & 248 & 16 & 6.49 \\
 2010.63 & 29 & 518 & 24 & 6.54 \\
 2011.60 & 37 & 479 & 30 & 6.47 \\
 2012.72 & 37 & 620 & 34 & 6.54 \\
 2013.71 & 27 & 423 & 41 & 6.58 \\
 2014.65 & 15 & 190 & 48 & 6.03 \\
 2014.89 & 38 & 385 & 51 & 6.40 \\
 2016.11 & 43 & 421 & 54 & 6.79 \\
 2017.23 & 19 & 279 & 59 & 6.61 \\
 2018.49 & 22 & 381 & & \\
 2019.23 & 34 & 195 & & \\
 2019.88 & 16 & 205 & & \\
 2020.79 & 26 & 193 & & \\
 2020.86 & 12 & 162 & & \\
 2021.09 & 14 & 230 & & \\
 \hline
 \end{tabular}
 \end{table}

 \begin{table}[hbt!]
 \caption{Magnitude 5 events from ~\citet{MIC_2019}.}
 \centering
 \begin{tabular}{c c c c}
 \hline
 Time & Event number & Magnitude & Sub-event of bigger event \\
 \hline
 2007.06 & 1 & 5.64 & Yes \\
 2007.08 & 2 & 5.91 & Yes \\
 2008.38 & 11 & 5.50 & Yes \\
 2009.16 & 14 & 5.50 & No \\
 2009.36 & 17 & 5.32 & Yes \\
 2010.63 & 25 & 5.76 & Yes \\
 2011.66 & 31 & 5.61 & Yes \\
 2011.66 & 32 & 5.32 & Yes \\
 2012.69 & 35 & 5.56 & Yes \\
 2013.74 & 42 & 5.71 & Yes \\
 2014.69 & 49 & 5.31 & Yes \\
 2014.93 & 52 & 5.39 & Yes \\
 2016.03 & 57 & 5.80 & Yes \\
 2017.13 & 60 & 5.43 & Yes \\
 2017.22 & 61 & 5.37 & Yes \\
 \hline
 \end{tabular}
 \end{table}

\newpage

\section*{Figure captions}

\begin{itemize}

\item Figure1. Details and smooth of the wavelet decomposition of a synthetics signal with period 500 days and duration of the slow slip event equal to 2 days (left), 5 days, 10 days, and 20 days (right).

\item Figure2. Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right). Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).

\item Figure3. Details and smooth of the wavelet decomposition of the longitudinal displacement recorded at GPS station PGC5.

\item Figure 4. GPS stations used in this study (black triangles). The black line represents the 40 km depth contour of the plate boundary model by ~\citet{PRE_2003}. The red triangles are the locations where we stack the GPS data. The small grey dots are all the tremor locations from the PNSN catalog.

\item Figure 5. Details and smooth of the wavelet decomposition of the detrended cumulative tremor count around the third northernmost location on Figure 3.

\item Figure 6. Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.

\item Figure 7. Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.

\item Figure 8. Top: Stacked 6th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 6th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.

\item Figure 9. Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the sum of the 6th, 7th and 8th level details of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.

\item Figure 10. ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition. The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.

\item Figure 11. Top: Stacked 5th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3.

\end {itemize}

\newpage

\section*{Figures}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_8.eps}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_8.eps}
\caption{Top: Stacked 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 8th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_7.eps}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_7.eps}
\caption{Top: Stacked 7th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the 7th level detail of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6.eps}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/GPS_longer_detail_6-7-8.eps}

\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/tremor_longer_detail_6-7-8.eps}
\caption{Top: Sum of the stacked 6th, 7th and 8th level details of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. Bottom: Opposite of the sum of the 6th, 7th and 8th level details of the cumulative tremor count in a 50 km radius of a given point for the same 16 locations.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/ROC_6-7-8.eps}
\caption{ROC curve for the sum of the 6th, 7th, and 8th level details of the wavelet decomposition. The red cross marks the true positive rate and the false positive rate obtained with the thresholds used to make Figure 9.}
\label{pngfiguresample}
\end{figure}

@@
